---
title: "Statistical Inference Project Part 1 - A Simulation Exercise"
author: "P.Y.Fong"
date: "12/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library (knitr)
opts_chunk$set(cache =F,
               echo = T,
               message=F,
               warning=F)
```

## Overview:  
Part 1 of this project investigates the exponential distribution in R and compares it with the Central Limit Theorem.  

### Simulation:  

The exponential distribution was simulated in R with *rexp(n, lambda)* where lambda is the rate parameter and *n* is the number of observations.  
The distribution of the averages of 40 exponentials was investigated across 1000 simulations.  
The mean of the exponential distribution is $\frac{1}{\lambda}$ and the standard deviation is also $\frac{1}{\lambda}$.  
lambda was set to 0.2 for all of the simulations.  

R Script for the simulation:
```{r simulation}
# Initialising simulation control variables and reproducible seed.  
numsim <- 1000
sampsize <- 40
lambda <- 0.2
set.seed(100)

# Store exp dist sim results in matrix with 1000 rows (simulations) and 40 columns (samples).  
matsim <- matrix(rexp(n = numsim * sampsize, rate = lambda), numsim, sampsize)

# Calculate the mean of each row of the sim. matrix.  
meansim <- rowMeans(matsim)
```  

Visualisation of the simulation mean distribution:
```{r Plots ,fig.cap= "Averages for Exponential Distribution (lambda= 0.2)"}
# Histogram
hist(meansim, breaks=30, prob=T, col='steel blue', xlim=c(2,8), ylim= c(0,.6), main="", xlab="")

# Distribution Averages Outline
lines(density(meansim),col= 'dark blue',lwd= 2)

# Theoretical mean
abline(v=1/lambda, col="red", lwd= 1)
abline(v=mean(meansim), col='black', lwd= 1)
text(x=5,y=0.6, " Theoretical Mean = 5 ", col = "red", adj = c(0, .5))
text(x=5,y=0.6, paste("Sample Mean=", round(mean(meansim),2)," "), col = "black", adj = c(1, .5))

# Theoretical density.
xfit <- seq(min(meansim), max(meansim), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sampsize)))
lines(xfit, yfit, pch=22, col="red", lty=2, lwd= 2)

legend('right', c("Simulation", "Theoretical"), lty=c(1,2), col=c("black", "red"))
```  

### Sample Versus Theoretical :  
The sample mean is `r round(mean(meansim),3)`, and sample variance is `r round(var(meansim),3)`,  
while the theoretical mean is $\frac{1}{\lambda}$ = `r round(1/lambda,3)` and theoretical variance is $\frac{1}{(n*{\sigma}^2)}$ = `r round(1/(40 *lambda^2),3)`.  

## Does the simulation approximate the Normal Distribution?:  
From figure.1  it can be seen that the simulation distribution approximates a Gaussian distribution. 

A quantile to quantile (q-q) plot also suggests a normal distribution.

```{r qqplot}
qqnorm(meansim); qqline(meansim)
```

